Information Closure Theory of Consciousness

### 2020-04-16 ICT_Frontiers_rev202001.tex
Updates from Overleaf

%% ============================== Useful packages ============================= %
\usepackage[dvipsnames,table,xcdraw]{xcolor}
\usepackage[round, sort]{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{svg}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{authblk}
\usepackage{float}
\usepackage{tikz}
\usepackage{lineno}
\usepackage{csquotes}

@@

		% Definition of informational closure
			\linelabel{line:r1q1a}Consider two processes, the environment process $(E_t)_{t \in \mathbb{N}}$ and the system's process $(Y_t)_{t \in \mathbb{N}}$ and let their interaction be described by the Bayesian network in Fig.~\ref{fig:SystemAndEnv}. Then, information flow $J_{t}$ from the environment $E$ to a system $S$ at time $t$ can be defined as the conditional mutual information $I$ between the current environment state $E_{t}$  and the future system state $Y_{t+1}$ given the current system state $Y_{t}$

@@

			\subsection{Informational Closure Does not Imply Causality}\label{sec:causality}
                A surprising (but well known) result from the definition of information flow $J_t(E\rightarrow Y)$ (Eq. \ref{eq:InformationFlow}) is that information flow does not indicate causal dependency from $E_t$ to $Y_{t+1}$ and from $Y_t$ to $Y_{t+1}$. Here we consider two scenarios, called "modelling" and "passive adaptation", which were already noted by \citet{BERTSCHINGER.2006}. In both scenarios, a process can form positive NTIC ($NTIC(E\rightarrow Y)> 0$) and informational closure ($J(E\rightarrow Y)=0$), however, via different causal dependencies.

@@

                However, in most of the realistic cases, the environment is partially observable to the system and, thus, the sensory process is usually not deterministic. Therefore, it is difficult for the system to be informationally closed and have higher NTIC.
                More importantly, we argue in Appendix \ref{sec:appendix} that whenever the environment has itself more predictable dynamics than the observations, there is potential for a process to achieve higher NTIC by modelling the environment than by copying the observations.

@@

        \noindent Note that, here we applied the same definitions of information flow (Eq.~\ref{eq:InformationFlow}) \linelabel{line:IC2CG}
        \begin{align}
            J_{t}(E \rightarrow Y )=I(Y_{t+1};E_t|Y_t)
        \end{align}
            to the system-environment dependency
            and the micro-macro scale dependency
        \begin{align}
            J_{t}(X \rightarrow Y )=I(Y_{t+1};X_t|Y_t)
        \end{align}
        even though the Bayesian graphs are different in the two scenarios. Both of these settings have already been used in the literature \citep[see][]{BERTSCHINGER.2006, PFANTE.2014}).\\\\

@@

        % ====================================
        % Refinement construction approach
        % ====================================
        % \begin{hypothesis}
        %         For any coarse-graining $Y$ of a stochastic process $X$, a refinement $S$ of $Y$ and the environment $E$ of $S$.  
        %         The \emph{content $C_t^{Content}(Y)$ of the consciousness of $Y$ at time $t$} is the state $y_t$ of $Y$ at time $t$ and \emph{the level $C_t^{Level}(Y)$ of the consciousness of $Y$} is the degree of NTIC of the C-process to the environment minus the information flow from the microscopic process $X$ to $Y$:
        %         \begin{align}
        %             C_t^{Content} &= y_t \label{eq:cContent}\\
        %             C_t^{Level} &= NTIC_t(E\rightarrow Y)-J_t(X \rightarrow Y) \label{eq:cLevel}
        %         \end{align}
        %         \end{hypothesis}
        % ====================================

        A concrete example in the context of neuroscience is that $X$ represents the microscopic scale of the universe, $S$ a cellular scale process in the neural system, $Y$ a more macroscopic process of the neural system coarse-grained from the cellular scale process $S$, and $E$ the environment which the cellular level process $S$ interacts with. \linelabel{line:neu-env} The environment $E$ may include other processes in the neural system, the sensors for perception and interoception, and external physical worlds.\\

@@

            \item[Implication 5.]
            Conscious processes can encode environmental information (by forming NTIC), however, be ignorant to part of the information of more microscopic processes (from Implication 3 and 4). This is in line with our conscious experience that information that every conscious percept provides represent rich and structured environmental states without involving all the information about microscopic activities.
        \end{description}

@@

    		% CC doesn't include all microscopic states
    		Informational closure can happen between scales of coarse-graining within a single system. Thus, a macroscopic NTIC process can be ignorant to its microscopic states. \ac{OurTheory} argues that human conscious contents do not reflect cellular scale activity because the conscious process which corresponds to a macroscopic NTIC process is informationally closed to the cellular scale in the human neural system. Further more, since NTIC processes are informationally closed, each of them can be considered as a reality. In the extreme case, when the information flow from its microscopic processes and environment to the informationally closed process is completely zero (Eq.~\ref{eq:informationflow2}), the future states of the process is only determined by its past states.

    		\linelabel{line:encode}Importantly, NTIC processes internalise the environmental dynamics in its states \citep[also see P.~4][]{BERTSCHINGER.2006}. This suggests that a NTIC process can be considered as a process that models the environmental dynamics.This implication fits well with some theories of consciousness (for example, world simulation metaphor ~\citep{revonsuo2006inner}). Note that \ac{OurTheory} doesn't assume that generative models are necessary for consciousness. The implication is a natural result of processes with NTIC.

@@

    	    According to Sec.~\ref{sec:cl} and Sec.~\ref{sec:cc}, an important implication from \ac{OurTheory} is that both conscious levels and conscious contents are associated with the state space of an NTIC process $Y$. A large state space of $Y$ contributes conscious levels through the mutual information $I(Y_{t+1};Y_{t})$ and also contributes richer conscious contents by providing more possible states of conscious processes. 
    	    \ac{OurTheory} therefore explains why, in normal physiological states, conscious levels and conscious contents are often positively correlated \citep{laureys2005neural}. This implication is also in line with the intuition in which consciousness is often associated with complex systems.

@@

            \subsubsection*{Processes that are not Informationally Closed}\label{sec:reflexive}
                % ------------------------------------
                % This section should follow:
                % 1. Processes that are not informationally closed is unconscious
                % 2. Some informationally not closed processes are passively driven by the environment (some examples)
                % 3. This may give readers the impression that processes passively driven by the environment are all not informationally closed
                % 4. This is a wrong impression because in a specific circumstance which the sensory processes are deterministic, passive adaptation can be informational closed.
                % 5. However, 4 is difficult to get because in realistic cases, the environment is often partially observable and this leads to sensory processes being non-deterministic. Therefore, ICT predict that in most of the realistic cases, passive adaptation (e.g., feed-forward network, video camera) is unconscious.
                % ------------------------------------
                \linelabel{line:notIC} 
                % To maximise NTIC, one can minimise the information flow $J_{t}(E \rightarrow Y )$ from the environment $E$ to the process $Y$ (Eq.~\ref{eq:InformationFlow},\ref{eq:nticObjective2}). When the information flow is zero, the process is informationally closed with respect to the environment. 
                The first scenario is built upon the assumption that sensor processes are non-deterministic \footnote{Non-deterministic sensor processes here means $H(\hat{e}_{t+1}|\hat{e}_t)>0$.} and the process dynamics are passively driven by the environmental inputs. Such processes cannot be informationally closed and are therefore unconscious.

                % When the sensor processes are not deterministic, if processes dynamics is passively driven by the environmental inputs the process is not informationally closed and, thus, unconscious.

                Reflexive behaviours \citep{casali2013theoretically} can be considered as an example of this scenario. In \ac{OurTheory}, if we can view reflexive behaviours as situations in which the internal state $Y_t$, which triggers reflexive action, is determined by the environment state $E_{t-1}$, overruling the influences from its own past $Y_{t-1}$. Such interpretation of reflexive behaviour from the viewpoint of \ac{OurTheory} naturally explains why reflexes involve less or no conscious experience of external stimuli.

        		\begin{figure}[H]
        			\includegraphics[width=0.8\textwidth]{Submission/Frontiers/Reflexive.pdf}
        			\caption{
        			    A diagram depicting the information flow in reflexive behaviours (shown by the red nodes and arrows) happening through the interaction between a process $Y$ and its environment $E$. In such situations, the internal state $Y_t$ is mostly dependent on the environment state $E_{t-1}$ but less on its past state $Y_{t-1}$. Therefore, the process $Y$ is not informational closed. As a  consequence, $Y$ is unable to form high NTIC and, therefore, remain less conscious or unconscious (see Fig.~\ref{fig:reflexive}). 
        			    }
        			\label{fig:reflexive}
        		\end{figure}

                The same principle can be applied to interpret blindsight \citep{humphrey1999history, humphrey1974vision, Humphrey1970} and procedural memory \citep{doyon2009contributions, ashby2010cortical} which are often considered as unconscious processes.
        		\linelabel{line:r2q6}Blindsight patients are able to track objects, avoid obstacles, and make above chance-level visual judgements with degraded or missing visual experience. However, in some cases, they may still preserve some forms of conscious experience (see \cite{overgaard2011visual, mazzi2016blind}). We argue that blindsight-guided actions are a result of 
        		% action outputs of blindsight can be directly guided by sensory inputs through 
        		stimulus-response mapping. The corresponding neural circuits are driven passively and therefore are not informationally closed. According to \ac{OurTheory} we therefore have no conscious experience of the visual stimuli.

@@

                Note that, not all processes that are driven by the environment (passive adaptation) are unconscious. When the sensor processes are deterministic, a system can still have positive NTIC and achieve informational closure via passive adaptation. \linelabel{line:feedforward}Therefore, some passive system (for example pure feedforward networks) can potentially produce consciousness.\footnote{Since a $n$-layer feedforward network is a system with $n$-step memory it is technically appropriate to use the $n$-step memory definition of NTIC, i.e. $NTIC^m_t(E\rightarrow Y):=I(Y_{t+1}:E_t,...,E_{t-n+1})-I(Y_{t+1}:E_t,...,E_{t-n+1}|Y_t)$ \citep{BERTSCHINGER.2006}, for such systems. In this case the notion of non-deterministic input processes should be generalised to input processes with $H(\hat{e}_t|\hat_{e}_{t-1},...,\hat{e}_{t-n})>0$.}

                For agents such as human beings, the environment is often informationally rich but only partially observable in such a way that the current sensory inputs are insufficient to predict the next and to form deterministic sensor processes. In this circumstance, the system cannot become informationally closed by passive adaptation (e.g., simply copying the sensory values to the system). ICT predicts that, in most of the realistic cases, processes with passive adaptation are unconscious. On the other hand, networks with recurrent loops employ information stored in their own past states have the potential to achieve higher NTIC by modelling the environment. If it turns out to be true that for every pure feed-forward network there are non-feed-forward systems achieving higher NTIC then ICT predicts that the latter systems achieve higher levels of consciousness. This implication would coincide with theories of consciousness emphasising the importance of recurrent circuits to consciousness \citep{lamme2006towards, edelman1992bright, tononi2008neural}.

                % ==== end ===

                % Finally, with a caveat, when the sensory dynamics is deterministic, a system can still have positive NTIC and achieve informational closure by simply copying the sensory values to its internal state (i.e. passive adaptation, see Sec.~\ref{sec:causality}). ICT will predict that, in this specific scenario, the systems can have some levels of consciousness. Although this specific scenario is uncommon in realistic circumstance an important implication is that pure feedforward networks can potentially produce consciousness. The level of consciousness for these networks is bounded by the sensory channel capacity and depth of the network. On the other hand, networks with recurrent loops employ information stored in their own past states have the potential to achieve higher NTIC by modelling the environment. If it turns out to be true that for every pure feed-forward network there are non-feed-forward systems achieving higher NTIC then ICT predicts that the latter systems achieve higher levels of consciousness.

@@

        % 		cannot produce consciousness because NTIC requires a form of memory I(Y t+1 ; Y t ) . Without memory, a network’s current state is entirely driven by the input the network receives from the environment without any inﬂuence from its own past states. Therefore, such a network is incapable of forming an NTIC. In contrast, a network with recurrent loops can maintain information about its own past states. This forms an information channel between the past and the future states of the network and, thus, makes the network capable of being informationally closed. This result coincides with theories of consciousness emphasising the importance of recurrent circuits to consciousness (Edelman, 1992; Lamme, 2006; Tononi & Koch, 2008).
        % 		than the because NTIC requires a form of memory $I(Y_{t+1};Y_{t})$. Without memory, a network's current state is entirely driven by the input the network receives from the environment without any influence from its own past states. Therefore, such a network is incapable of forming an NTIC. In contrast,

            \subsubsection*{Processes that are Trivially Closed}
                The second scenario is that when encoded information in a process is trivial, i.e. no mutual information between the process states and its environment states $I(Y_{t+1};E_{t})$ (Eq.~\ref{eq:nticObjective2}), this leads to non-positive NTIC. In such cases, the process is considered to be unconscious. This implies that an isolated process which is informationally closed is insufficient to be conscious. 
                This mathematical property of \ac{OurTheory} 
                is relevant for 
                % provides a natural and intuitive (but only partial, see the current challenge in  Sec.~\ref{sec:Limitation and Future work}) solution to 
                dealing with 
                the boundary and the individuality problems of consciousness\footnote{The boundary problem of consciousness refers to identifying physical boundaries of conscious processes and the individuality problem of consciousness refers to identifying individual consciousnesses in the universe.}
                \citep{Raymont2006-RAYUOC}. Consider a NTIC process $Y$ and an isolated informationally closed process $\hat{Y}$ with only trivial information. Adding $\hat{Y}$ to $Y$ can still maintain informational closure but does not increase non-trivial information, i.e.,  consciousness is unaffected.

@@

                This implies that isolated processes with trivial information do not contribute consciousness and should be considered being outside the information boundary of the conscious processing. 
                % (for more details of the boundary detection procedure, see \cite{krakauer2014information}). 
                This property also implies that consciousnesses do not emerge from just aggregating informationally closed (isolated) processes which contain trivial information.
                In the future we hope to adapt the procedures for boundary detection proposed in \citet{krakauer2014information,krakauer2020information} to ICT.

@@

		    Previous consciousness research has identified a number of diverse cognitive processes often accompanied by conscious experience. \ac{OurTheory} provides an integrated account for the reason why these processes involve conscious experience. As mentioned above, an NTIC process can be seen as an internal simulation engine for the agent-environmental interactions \citep{BERTSCHINGER.2006}. Therefore, information encoded in NTIC processes is essential for several cognitive processes.

@@

            Second, as a simulation engine, with a given initial state, an NTIC process can self-evolve and simulate the environmental transitions. Cognitive functions involving simulations are expected to involve NTIC processes. Consequently, mental simulation, imagination, computing alternative realities, and generating counterfactuals often come with conscious experience.

@@

    	    Finally, for agents adapting to complex environments (e.g., human being), any state of the NTIC process can be seen as an integration of high dimensional information. To accurately encode information about the complex environmental states and transitions, the NTIC process requires knowledge about the complex causal dependencies involved in the environment. Therefore, cognitive functions requiring large scale integration are likely to involve NTIC processes and accompanied by conscious experience. % Furthermore, the high dimensional integrated information is also critical for an agent to learn new behaviours to achieve flexibility and the strategic behavioural controls and decision-making.

@@

        % ---------------------------------------------------------------------------- %
        %        Theories about multilevel views on consciousness and cognition        %
        % ---------------------------------------------------------------------------- %
        \subsection{Multilevel Views on Consciousness and Cognition}\label{sec:MultiLevelView}
    		\ac{OurTheory} proposes that conscious processes can occur at any scale of coarse-graining which forms NTIC within a system. This suggests that the scale of coarse-graining is critical for searching and identifying the information corresponding to consciousness. A few versions of multilevel views on consciousness have previously been (explicitly or implicitly) proposed. To our knowledge, Pennartz's neurorepresentational theory (also called Neurorepresentationalism,~\citep{pennartz2018consciousness,pennartz2015brain}) is the proposal closest to the multilevel view of \ac{OurTheory}. Similar to Neurorepresentationalism, the concept of levels in \ac{OurTheory} is also relevant to Marr's level of analysis~\citep{marr1982vision, pennartz2015brain, pennartz2018consciousness}. 
    		However, \ac{OurTheory} suggests that coarse-graining is necessary only when the microscopic processes are stochastic (e.g. the neural system). An NTIC process can be formed in a noise-free deterministic system without coarse-graining. According to \ac{OurTheory}, this NTIC process is sufficient to be conscious. 
    		Another fundamental difference between \ac{OurTheory} and Neurorepresentationalism is that Neurorepresentationalism takes functionalist perspective and suggests consciousness should serve high-level world-modelling and makes a best guess about the interaction between the body and the environment. 
    		%Neurorepresentationalism also suggests conscious experience is associated with integrated representation for multimodal and situational information.
    		However, \ac{OurTheory} is grounded by non-functional informational hypothesis. Therefore, \ac{OurTheory} provides a more fundamental explanation for the scale problem of consciousness.

@@

    		Here, we want to make clear that the "level" in \ac{OurTheory} refers to the \textbf{scales of coarse-graining} instead of the "level" for the cortical anatomy or sensory processing. It is important to note that the coarse-graining direction is an orthogonal dimension irrespective of the level of anatomy or the level of information processing hierarchy in the neural system (see Fig.~\ref{fig:hierarchy}). Because ILT focuses on the levels of the sensory processing hierarchy and \ac{OurTheory} focus on informational closure among the levels of coarse-graining, the two theories are fundamentally different.

@@

		    \textbf{The concept of "integration"}: In IIT, integrated information is one of the core concept of defining conscious individuals. In the current paper, we did not include the notion of integrated information within \ac{OurTheory}. This, however, results in one of the current weaknesses of \ac{OurTheory} that it lacks the ability to individuate NTIC processes in some extreme cases (i.e., the problem of individuality). We discussed this weaknesses in Sec.~\ref{sec:Limitation and Future work}.

@@

    		% fill the gaps
    		So, eventually, why are some predictive information conscious and some are not? \ac{OurTheory} predicts that only the predictions generated from mechanisms involving the NTIC process are conscious. Note that predictive processes are not necessary to involve NTIC processes. A predictive process can make prediction about the future state of its environment based on the current sensory inputs. In this case, the the process is not informationally closed and could not be conscious.

@@

        % Can't solve the hard problem
        It's important to clarify that \ac{OurTheory} does not intend to completely solve the hard problems of consciousness \citep{chalmers1995facing}. Knowing the state of a conscious process does not allow us to answer "What is it like to be in this state of this process" \citep{nagel1974like}. Instead, \ac{OurTheory} focuses more on bridging consciousness and the physical world using information theory as a common language in between.

>>>>

%% ============================== Useful packages ============================= %
\usepackage[dvipsnames,table,xcdraw]{xcolor}
\usepackage[round, sort]{natbib}
\usepackage{nameref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{svg}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{authblk}
\usepackage{float}
\usepackage{tikz}
\usepackage{lineno}
\usepackage{csquotes}

@@

		% Definition of informational closure
			\linelabel{line:r1q1a}Consider two processes, the environment process $(E_t)_{t \in \mathbb{N}}$ and the system's process $(Y_t)_{t \in \mathbb{N}}$ and let their interaction be described by the Bayesian network with the sensor channel $\hat{e}_{t}$ and the action $\hat{y}_{t}$ channel in Fig.~\ref{fig:SystemAndEnv}. Then, information flow $J_{t}$ from the environment $E$ to a system $S$ at time $t$ can be defined as the conditional mutual information $I$ between the current environment state $E_{t}$  and the future system state $Y_{t+1}$ given the current system state $Y_{t}$

@@

			\subsection{Informational Closure Does not Imply Causality}\label{sec:causality}
                A surprising result from the definition of information flow $J_t(E\rightarrow Y)$ (Eq.~\ref{eq:InformationFlow}) is that information flow does not indicate causal dependency from $E_t$ to $Y_{t+1}$ and from $Y_t$ to $Y_{t+1}$. Here we consider two scenarios, called \textit{modelling} and \textit{passive adaptation}, which were already noted by \citet{BERTSCHINGER.2006}. In both scenarios, a process can form positive NTIC ($NTIC(E\rightarrow Y)> 0$) and informational closure ($J(E\rightarrow Y)=0$), however, via different causal dependencies.

@@

                However, in most of the realistic cases, the environment is partially observable to the system and, thus, the sensory process is usually not deterministic. Therefore, it is difficult for the system to be informationally closed and have higher NTIC.
                More importantly, we argue in Appendix that whenever the environment has itself more predictable dynamics than the observations, there is potential for a process to achieve higher NTIC by modelling the environment than by copying the observations.

@@

        \noindent Note that, here we applied the same definitions of information flow (Eq.~\ref{eq:InformationFlow}) \linelabel{line:IC2CG}
        \begin{align}
            J_{t}(E \rightarrow Y )=I(Y_{t+1};E_t|Y_t)
        \end{align}
            to the system-environment dependency
            and the micro-macro scale dependency
        \begin{align}
            J_{t}(X \rightarrow Y )=I(Y_{t+1};X_t|Y_t)
        \end{align}
        Even though the Bayesian graphs are different in the two scenarios. Both of these settings have already been used in the literature \citep[see][]{BERTSCHINGER.2006, PFANTE.2014}).\\

@@

        A concrete example in the context of neuroscience is that $X$ represents the microscopic scale of the universe, $S$ a cellular scale process in the neural system, $Y$ a more macroscopic process of the neural system coarse-grained from the cellular scale process $S$, and $E$ the environment which the cellular level process $S$ interacts with. \linelabel{line:neu-env} The environment $E$ may include other processes in the neural system, the sensors for perception and interoception, and external physical worlds.

@@

            \item[Implication 5.]
            Conscious processes can model environmental information (by forming NTIC), however, be ignorant to part of the information of more microscopic processes (from Implication 3 and 4). This is in line with our conscious experience that information that every conscious percept provides represent rich and structured environmental states without involving all the information about microscopic activities.
        \end{description}

@@

    		% CC doesn't include all microscopic states
    		Informational closure can happen between scales of coarse-graining within a single system. Thus, a macroscopic NTIC process can be ignorant to its microscopic states. \ac{OurTheory} argues that human conscious contents do not reflect cellular scale activity because the conscious process which corresponds to a macroscopic NTIC process is informationally closed to the cellular scale in the human neural system. Further more, since NTIC processes are informationally closed, each of them can be considered as a reality. When the information flow from its microscopic processes and environment to the informationally closed process is completely zero (Eq.~\ref{eq:informationflow2}), the future states of the process can entirely be determined by its past states.

    		\linelabel{line:encode}Importantly, in most of the realistic cases, NTIC processes internalise the environmental dynamics in its states (see Sec. \ref{sec:causality} and also \cite{BERTSCHINGER.2006}). This suggests that a NTIC process can be considered as a process that models the environmental dynamics. This implication fits well with several theories of consciousness (for example, world simulation metaphor ~\citep{revonsuo2006inner}). Note that \ac{OurTheory} does not assume that generative models are necessary for consciousness. The implication is a natural result of processes with NTIC.

@@

    	    According to Sec.~\ref{sec:cl} and Sec.~\ref{sec:cc}, an important implication from \ac{OurTheory} is that both conscious levels and conscious contents are associated with the state space of an NTIC process $Y$. A larger state space of $Y$ contributes conscious levels through the mutual information $I(Y_{t+1};Y_{t})$ and also contributes richer conscious contents by providing more possible states of conscious processes. 
    	    \ac{OurTheory} therefore explains why, in normal physiological states, conscious levels and conscious contents are often positively correlated \citep{laureys2005neural}. This implication is also in line with the intuition in which consciousness is often associated with complex systems.

@@

            \subsubsection*{Processes that are not Informationally Closed}\label{sec:reflexive}
                % ------------------------------------
                % This section should follow:
                % 1. Processes that are not informationally closed is unconscious
                % 2. Some informationally not closed processes are passively driven by the environment (some examples)
                % 3. This may give readers the impression that processes passively driven by the environment are all not informationally closed
                % 4. This is a wrong impression because in a specific circumstance which the sensory processes are deterministic, passive adaptation can be informational closed.
                % 5. However, 4 is difficult to get because in realistic cases, the environment is often partially observable and this leads to sensory processes being non-deterministic. Therefore, ICT predict that in most of the realistic cases, passive adaptation (e.g., feed-forward network, video camera) is unconscious.
                % ------------------------------------
                \linelabel{line:notIC} 
                % To maximise NTIC, one can minimise the information flow $J_{t}(E \rightarrow Y )$ from the environment $E$ to the process $Y$ (Eq.~\ref{eq:InformationFlow},\ref{eq:nticObjective2}). When the information flow is zero, the process is informationally closed with respect to the environment. 
                The first scenario is built upon the assumption that sensor processes are non-deterministic \footnote{Non-deterministic sensor processes here means $H(\hat{e}_{t+1}|\hat{e}_t)>0$.} and the process dynamics are passively driven by the environmental inputs. Such processes cannot be informationally closed and are, therefore, unconscious.

                Reflexive behaviours \citep{casali2013theoretically} can be considered as an example of this scenario. In \ac{OurTheory}, if we can view reflexive behaviours as situations in which (Fig.~\ref{fig:reflexive}) the internal state $Y_t$, which triggers reflexive action $\hat{y}_t$, is determined by the environment state $E_{t-1}$, overruling the influences from its own past $Y_{t-1}$. Such interpretation of reflexive behaviour from the viewpoint of \ac{OurTheory} naturally explains why reflexes involve less or no conscious experience of external stimuli.

        		\begin{figure}[H]
        			\includegraphics[width=0.8\textwidth]{Submission/Frontiers/Reflexive.pdf}
        			\caption{
        			    A diagram depicting the information flow in reflexive behaviours (shown by the red nodes and arrows) happening through the interaction between a process $Y$ and its environment $E$. When the sensor process $\hat{e}_t$ is  non-deterministic and the internal state $Y_t$ is mostly dependent on the sensor state $\hat{e}_t$ driven by the environment $E_{t-1}$ but less on its past state $Y_{t-1}$. As a consequence, $Y$ is unable to form informational closure and, therefore, remain unconscious.}
        			\label{fig:reflexive}
        		\end{figure}

                The same principle can be applied to interpret blindsight \citep{humphrey1999history, humphrey1974vision, Humphrey1970} and procedural memory \citep{doyon2009contributions, ashby2010cortical} which are often considered as unconscious processes.
        		\linelabel{line:r2q6}Blindsight patients are able to track objects, avoid obstacles, and make above chance-level visual judgements with degraded or missing visual experience.(However, in some cases, they may still preserve some forms of conscious experience; See \cite{overgaard2011visual, mazzi2016blind}). We argue that blindsight-guided actions are a result of 
        		stimulus-response mapping. The corresponding neural circuits are driven passively and therefore are not informationally closed. According to \ac{OurTheory} we therefore have no conscious visual experience of visual the  stimuli.

@@

                Note that, not all processes that are driven by the environment (passive adaptation) are unconscious. As mentioned in Sec.~\ref{sec:causality}, when the sensor processes are deterministic, a system can still have positive NTIC and achieve informational closure via passive adaptation. \linelabel{line:feedforward}Therefore, some passive system (for example pure feedforward networks) can potentially be conscious.\footnote{Since a $n$-layer feedforward network is a system with $n$-step memory it is technically appropriate to use the $n$-step memory definition of NTIC, i.e.~$NTIC^m_t(E\rightarrow Y):=I(Y_{t+1}:E_t,\dots,E_{t-n+1})-I(Y_{t+1}:E_t,\dots,E_{t-n+1}|Y_t)$ \citep{BERTSCHINGER.2006}, for such systems. In this case the notion of non-deterministic input processes should be generalised to input processes with $H(\hat{e}_t|\hat{e}_{t-1},\dots,\hat{e}_{t-n})>0$.}

                For agents such as human beings, the environment is often informationally rich but only partially observable in such a way that the current sensory inputs are insufficient to predict the next and to form deterministic sensor processes. In this circumstance, the system cannot become informationally closed by passive adaptation (e.g., simply copying the sensory values to the system). ICT predicts that, in most of the realistic cases, processes with passive adaptation are unconscious. On the other hand, networks with recurrent loops employ information stored in their own past states have the potential to achieve higher NTIC by modelling the environment. If it turns out to be true that for every pure feed-forward network there are non-feed-forward systems achieving higher NTIC then ICT predicts that the latter systems achieve higher levels of consciousness. This implication would coincide with theories of consciousness emphasising the importance of recurrent circuits to consciousness \citep{lamme2006towards, edelman1992bright, tononi2008neural}.                
                % Finally, with a caveat, when the sensory dynamics is deterministic, a system can still have positive NTIC and achieve informational closure by simply copying the sensory values to its internal state (i.e. passive adaptation, see Sec.~\ref{sec:causality}). ICT will predict that, in this specific scenario, the systems can have some levels of consciousness. Although this specific scenario is uncommon in realistic circumstance an important implication is that pure feedforward networks can potentially produce consciousness. The level of consciousness for these networks is bounded by the sensory channel capacity and depth of the network. On the other hand, networks with recurrent loops employ information stored in their own past states have the potential to achieve higher NTIC by modelling the environment. If it turns out to be true that for every pure feed-forward network there are non-feed-forward systems achieving higher NTIC then ICT predicts that the latter systems achieve higher levels of consciousness.

@@

        % 		cannot produce consciousness because NTIC requires a form of memory I(Y t+1 ; Y t ) . Without memory, a network’s current state is entirely driven by the input the network receives from the environment without any inﬂuence from its own past states. Therefore, such a network is incapable of forming an NTIC. In contrast, a network with recurrent loops can maintain information about its own past states. This forms an information channel between the past and the future states of the network and, thus, makes the network capable of being informationally closed. This result coincides with theories of consciousness emphasising the importance of recurrent circuits to consciousness (Edelman, 1992; Lamme, 2006; Tononi & Koch, 2008).
        % 		than the because NTIC requires a form of memory $I(Y_{t+1};Y_{t})$. Without memory, a network's current state is entirely driven by the input the network receives from the environment without any influence from its own past states. Therefore, such a network is incapable of forming an NTIC. In contrast, 
            \subsubsection*{Processes that are Trivially Closed}
                The second scenario is that when encoded information in a process is trivial, i.e. no mutual information between the process states and its environment states $I(Y_{t+1};E_{t})$ (Eq.~\ref{eq:nticObjective2}), this leads to non-positive NTIC. In such cases, the process is considered to be unconscious. This implies that an isolated process which is informationally closed is insufficient to be conscious. 
                This mathematical property of \ac{OurTheory} 
                is relevant for 
                % provides a natural and intuitive (but only partial, see the current challenge in  Sec.~\ref{sec:Limitation and Future work}) solution to 
                dealing with 
                the boundary and the individuality problems of consciousness\footnote{The boundary problem of consciousness refers to identifying physical boundaries of conscious processes and the individuality problem of consciousness refers to identifying individual consciousnesses in the universe.}
                \citep{Raymont2006-RAYUOC}. Consider a NTIC process $Y$ and an isolated informationally closed process $\hat{Y}$ with only trivial information. Adding $\hat{Y}$ to $Y$ can still maintain informational closure but does not increase non-trivial information, i.e.,  consciousness is unaffected.

@@

                This implies that isolated processes with trivial information do not contribute consciousness and should be considered being outside the informational boundary of the conscious processing. 
                % (for more details of the boundary detection procedure, see \cite{krakauer2014information}). 
                This property also implies that consciousnesses do not emerge from just aggregating informationally closed (isolated) processes which contain trivial information.
                In the future we hope to adapt the procedures for boundary detection proposed in \citet{krakauer2014information,krakauer2020information} to ICT.

@@

		    Previous consciousness research has identified a number of diverse cognitive processes often accompanied by conscious experience. \ac{OurTheory} provides an integrated account for the reason why these processes involve conscious experience. As mentioned above, an NTIC process can be seen as an internal modelling engine for the agent-environmental interactions \citep{BERTSCHINGER.2006}. Therefore, information encoded in NTIC processes is essential for several cognitive processes.

@@

            Second, as a modelling engine, with a given initial state, an NTIC process can self-evolve and simulate the environmental transitions. Cognitive functions involving simulations are expected to involve NTIC processes. Consequently, mental simulation, imagination, computing alternative realities, and generating counterfactuals often come with conscious experience.

@@

    	    Finally, for agents adapting to complex environments (e.g., human being), any state of the NTIC process can be seen as an integration of high dimensional information. To accurately encode information about the complex environmental states and transitions, the NTIC process requires knowledge about the complex causal dependencies involved in the environment. Therefore, cognitive functions requiring larger scale integration are likely to involve NTIC processes and accompanied by conscious experience. % Furthermore, the high dimensional integrated information is also critical for an agent to learn new behaviours to achieve flexibility and the strategic behavioural controls and decision-making.

@@

        % ---------------------------------------------------------------------------- %
        %        Theories about multilevel views on consciousness and cognition        %
        % ---------------------------------------------------------------------------- %
        \subsection{Multilevel Views on Consciousness and Cognition}\label{sec:MultiLevelView}
    		\ac{OurTheory} proposes that conscious processes can occur at any scale of coarse-graining which forms NTIC within a system. This suggests that the scale of coarse-graining is critical for searching and identifying the information corresponding to consciousness. A few versions of multilevel views on consciousness have previously been (explicitly or implicitly) proposed. To our knowledge, Pennartz's neurorepresentational theory (also called Neurorepresentationalism,~\citep{pennartz2018consciousness,pennartz2015brain}) is the proposal closest to the multilevel view of \ac{OurTheory}. Similar to Neurorepresentationalism, the concept of levels in \ac{OurTheory} is relevant to Marr's level of analysis~\citep{marr1982vision, pennartz2015brain, pennartz2018consciousness}. 
    		However, \ac{OurTheory} suggests that coarse-graining is necessary only when the microscopic processes are not NTIC (e.g. the neural system). An NTIC process can be formed in a noise-free deterministic system without coarse-graining. According to \ac{OurTheory}, this NTIC process is sufficient to be conscious. 
    		Another fundamental difference between \ac{OurTheory} and Neurorepresentationalism is that Neurorepresentationalism takes functionalist perspective and suggests consciousness should serve high-level world-modelling and makes a best guess about the interaction between the body and the environment. 
    		%Neurorepresentationalism also suggests conscious experience is associated with integrated representation for multimodal and situational information.
    		However, \ac{OurTheory} is grounded by non-functional informational hypothesis. Therefore, \ac{OurTheory} provides a non-functional and fundamental explanation for the scale problem of consciousness.

@@

    		Here, we want to make clear that the "level" in \ac{OurTheory} refers to the \textbf{scale of coarse-graining} instead of the "level" for the cortical anatomy or sensory processing. It is important to note that the coarse-graining direction is an orthogonal dimension irrespective of the level of anatomy or the level of information processing hierarchy in the neural system (see Fig.~\ref{fig:hierarchy}). Because ILT focuses on the levels of the sensory processing hierarchy and \ac{OurTheory} focus on informational closure among the levels of coarse-graining, the two theories are fundamentally different.

@@

		    \textbf{The concept of "integration"}: In IIT, integrated information is one of the core concept of defining conscious individuals. In the current paper, we did not include the notion of integrated information within \ac{OurTheory}. This, however, results in one of the current weaknesses of \ac{OurTheory} that it lacks the ability to individuate NTIC processes in some cases (i.e., the problem of individuality). We discussed this weaknesses in Sec.~\ref{sec:Limitation and Future work}.

@@

    		% fill the gaps
    		So, eventually, why are some predictive information conscious and some are not? \ac{OurTheory} predicts that only the predictions generated from mechanisms involving the NTIC process are conscious. Note that predictive processes are not necessary to involve NTIC processes. A predictive process can make prediction about the future state of its environment solely based on the current sensor states when the current sensor states and future sensor states have positive mutual information. However, this is not sufficient for a process to be informationally closed and, therefore, be conscious.

@@

        % Can't solve the hard problem
        It's important to clarify that \ac{OurTheory} does not intend to  solve the hard problems of consciousness \citep{chalmers1995facing}. Knowing the state of a conscious process does not allow us to answer "What is it like to be in this state of this process" \citep{nagel1974like}. Instead, \ac{OurTheory} focuses more on bridging consciousness and the physical world using information theory as a common language in between.
